---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r loadData, echo=TRUE}

setwd("D:/RepData_PeerAssessment1/")
steps <- read.csv("activity.csv")

steps$date <- as.Date(as.character(steps$date))


```

## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day
```{r histogram_total_steps, echo=TRUE}

library(dplyr)
library(xtable)

steps_total <- steps %>%
  group_by(date) %>%
  summarize(sum = sum(steps))

hist(steps_total$sum, main = "Histogram of Total Number of Daily Steps", xlab = "Total Steps")

```

The mean total number of steps taken per day
```{r meanMedian_totalSteps, echo=TRUE}

mean_TS <- mean(steps_total$sum, na.rm = TRUE)
mean_TS

median_TS <- median(steps_total$sum, na.rm = TRUE)
median_TS

```

The mean total number of steps taken per day is `r format(round(mean_TS, 2), nsmall=2)`. The median total number of steps taken per day is `r format(round(median_TS,2), nsmall=2)`.

## What is the average daily activity pattern?
Time series plot of 5-minute interval and average number of steps taken
```{r avgdailyactivity, echo=TRUE}

library(ggplot2)

s_int <- steps %>%
  group_by(interval) %>%
  summarize(avg_steps = mean(steps, na.rm=TRUE))

ts_int <- ggplot(data=s_int, aes(x=interval, y=avg_steps))  + geom_line() + ggtitle("Average Steps Taken by Interval Across All Days") +xlab("Interval") + ylab("Average Steps")

ts_int

```

Five-minute interval, on average across all days, with maximum number of steps
```{r maxStepsInt, echo=TRUE}

max_int <- filter(s_int, avg_steps == max(s_int$avg_steps))

```
Interval `r max_int$interval`, with `r round(max_int$avg_steps,2), nsmall=2), steps, has the maximum number of steps, on average across all days in the dataset.


## Imputing missing values
```{r missingVals, echo=TRUE}

missing_count <- sum(is.na(steps$steps))

missing_count

```
There are `r missing_count` missing values in the dataset. 

Fill in missing values with 5-minute interval mean, and create new dataset with filled in missing values.
```{r fillMissing_vals, echo=TRUE}
steps_mi <- steps %>%
    group_by(interval) %>%
    mutate(meanInt = mean(steps, na.rm=TRUE))

steps_mi$steps <- ifelse(is.na(steps_mi$steps) == TRUE, steps_mi$meanInt, steps_mi$steps)

steps_mi1 <- data.frame(steps_mi[c(1:3)])


```

Histogram of total number of steps taken each day, with imputed missing values. 
```{r hist_total_nomv, echo=TRUE}

steps_mi1.sum <- steps_mi1 %>%
                      group_by(date) %>%
                      summarize(sum = sum(steps))


hist(steps_mi1.sum$sum, main = "Histogram of Total Number of Daily Steps (Imputed Missing Values)", xlab = "Total Steps")

mean_IV_TS <- mean(steps_mi1.sum$sum)
mean_IV_TS

median_IV_TS <- median(steps_mi1.sum$sum)
median_IV_TS

mean_var <- mean_TS - mean_IV_TS

median_var <- median_TS - median_IV_TS

```

The mean and median total number of steps taken per day with imputed missing values are:

1. mean: `r format(round(mean_IV_TS, 2), nsmall=2)`
2. median: `r format(round(median_IV_TS, 2), nsmall=2)`

The mean and median total number of steps taken per day without imputed missing values is `r format(round(mean_TS, 2), nsmall=2)` and `r format(round(median_TS,2), nsmall=2)`, respectively. 

The difference between the mean without imputed values and the mean with imputed values is `r format(round(mean_var,2), nsmall=2)`. 

The difference between the median without imputed values and the median with imputed values is `r format(round(median_var,2), nsmall=2)`. 


## Are there differences in activity patterns between weekdays and weekends?

```{r activitypatterns, echo=TRUE, results = "asis"}

steps_mi1$weekday <- weekdays(as.Date(steps_mi1$date))

steps_mi1$day_type <- gsub('Saturday|Sunday', 'weekend',steps_mi1$weekday )

steps_mi1$day_type1 <- gsub('Monday|Tuesday|Wednesday|Thursday|Friday', 'weekday',steps_mi1$day_type )

stepsmi1_wd <- steps_mi1 %>% 
          group_by(day_type1, interval) %>%
          summarize(meanint1 = mean(steps))

g01 <- ggplot(stepsmi1_wd, aes(interval, meanint1))
p01 <- g01  + facet_wrap(~ day_type1, nrow=2, ncol=1) + 
  geom_line(color="brown") + ylab("Number of Steps") + 
  labs(title = "Total Steps by Weekday Type") 

p01

```


